A unified pseudo-$C_\ell$ framework

### 2018-08-31 doc/Paper/main.tex

          \item {\bf Signal:} we generate maps of the galaxy overdensity $\delta$ and of the shear field $(\gamma_1,\gamma_2)$ for a redshift bin centred at $z\simeq1$ with a width $\Delta z\sim0.1$ and Gaussian photo-$z$ tails with width $\sigma_z=0.06$. For $\delta_g$ we assume unit bias $b(z)=1$ and no contributions from redshift-space distortions or magnification. For $\gamma$, we assume no contribution from intrinsic alignments. The maps are generated as Gaussian realizations of the corresponding angular power spectra computed with the Core Cosmology Library\footnote{\url{https://github.com/LSSTDESC/CCL/pulls}} for cosmological parameters compatible with the 2015 Planck measurements~\citet{2016A&A...594A..13P}. These power spectra are shown in Fig. \ref{fig:lss_cls}. Sample realizations of $\delta$ and $\gamma$ can be seen in Figures \ref{fig:lss_mocks_sph} and \ref{fig:lss_mocks_flat} for the curved-sky and flat-sky cases respectively.
          \item {\bf Mask:} for the curved-sky data, we use a mask built as a combination of the sky coverage output by the LSST {\tt OpSim} database~\citet{2014SPIE.9150E..15D} and a more conservative galactic cut based on the reddening map of \citet{1998wfsc.conf..297S}. To add an extra layer of complexity, we drill 100 additional 1-degree holes on the resulting combined mask. We also explore the improvement stemming from the down-weighting of higher-noise in an inverse-variance way in the presence of inhomogeneous noise (see description below), which is the reason why the default version of this mask is not binary, but traces the inverse of the noise variance (see top left panel of Fig. \ref{fig:lss_mocks_sph}). For the flat-sky data, we use a high-resolution mask constructed from the bright-object mask used for the first public data release from the HSC collaboration~\citep{2018PASJ...70S...4A}. The chosen footprint corresponds to the VVDS field, covers approximately 20 deg$^2$ and contains structure on a wide range of scales. As can be seen in the top-left panel of Figure~\ref{fig:lss_mocks_flat}, the mask is not binary, and its value at each pixel correspond to the pixel's fractional masked area.

@@

    \subsection{Curved-sky large-scale structure validation}\label{ssec:validation.lss_full}
      \begin{figure*}
        \centering
        \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph}
        \includegraphics[width=0.99\columnwidth]{val_covar_lss_sph}
        \caption{{\sl Left panel:} average over all simulations in our curved-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines). The analytical prediction for the input signal power spectra are shown as darker dashed lines, and lie mostly under the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$). The null power spectra ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$) are compatible with zero. The middle subplot shows the mean residuals with respect to the input power spectra normalized by the 1$\sigma$ error over the mean (i.e. the scatter over simulations divided by $\sqrt{N_{\rm sim}}$). The residuals are scattered around zero and their fluctuations are well within those expected under the assumption of Gaussian statistics given 1000 independent realizations (see Fig. \ref{fig:lss_val_chi2}). For comparison, the lower panel show the mean relative difference for the power spectra of contaminated maps without contaminant deprojection. {\sl Right panel:} correlation matrix ($C_{ij}/\sqrt{C_{ii}C_{jj}}$) of all measured power spectra estimated from 1000 realizations.} \label{fig:lss_val_cls}
      \end{figure*}
      \begin{figure*}
        \centering
        \includegraphics[width=0.80\textwidth]{val_chi2_lss_sph}
        \caption{Distribution of the $\chi^2$ values found for each of the 1000 simulations in our curved-sky LSS validation suite (blue histograms). Results are shown for all possible cross-correlations, labeled in the top right corner of each subplot. The simulated distributions are compared with the expected $\chi^2$ distribution (solid lines) assuming Gaussian statistics for $N_{\rm dof}=681$ degrees of freedom (denoted by the dashed vertical lines). The dot-dashed vertical lines mark the $\chi^2$ value found for the mean residual with respect to the input power spectrum (see middle panel of the left panel in Fig. \ref{fig:lss_val_cls}). The associated PTE values are $\gtrsim0.1$, and thus no statistically significant bias is found by our validation suite.} \label{fig:lss_val_chi2}
      \end{figure*}
      We generate 1000 realizations of contaminated maps of the galaxy overdensity $\delta$ and cosmic shear $\gamma$ using the models described in Section \ref{sssec:validation.suite.lss} with resolution $N_{\rm side}=1024$. In our fiducial run, we included all contaminants, and used a non-binary mask that traces the shot-noise fluctuations due to the presence of stars and depth variations. For each simulation, we use \nmt~to compute all six possible cross-correlations between the $\delta$ and the $E$ and $B$ modes of $\gamma$ (which we will label $\gamma_{E,B}$), including map-level contaminant deprojection and removal of the noise and deprojection bias in the power spectrum. We use band-powers defined by averaging over groups of $\Delta\ell=3$ multipoles. The typical running time for one realization is $\sim 600$s using 32 \texttt{OpenMP} threads. Subsequent runs are typically faster since we cache the coupling matrices, which are the slowest step in our power-spectra estimation. More details about timing and scalability of \nmt~can be found in Appendix \ref{app:scalability}.

@@

      The above figure was created using $E/B$ mode purification. Disabling this option has no impact on the bias of the estimates, but increases its variance as we show in the Figure \ref{fig:cmb_val_sigma}. We plot diagonal error bars with (red) and without (blue) $B$-mode purification for the $BB$ (solid), $EB$ (dashed) and $EE$ (dot-dashed) power spectra. Although there is no practical improvement in the error bars for the $EE$ spectrum (there is even a slight increase due to the loss of the ambiguous $E$ modes to purification), and the improvement is only mild for the $EB$ case, very significant gains can be obtained for the $BB$ power spectrum, reducing uncertainties by up to a factor of $\sim10$ on large scales.
      \begin{figure*}
        \centering
        \includegraphics[width=0.99\columnwidth]{val_cl_lss_flat}
        \includegraphics[width=0.99\columnwidth]{val_cl_cmb_flat}
        \caption{{\sl Left and Right:} the analogues of the left panels of Figures \ref{fig:lss_val_cls} and \ref{fig:cmb_cls} for the flat-sky LSS and CMB validation sets respectively.} \label{fig:val_cls_flat}
      \end{figure*}

@@

      \subsection{Pixelization and signal band-limits}\label{app:shts.blim}
        SHT transforms as written in the Eq.~\ref{eq:sht1} are, strictly speaking, defined on a continuous field on the sky and results in an infinite number of spherical harmonic coefficients. In practice, we have to always deal with discretized sky maps with a finite number of pixels at which the field is sampled, and consequently use a finite number of spherical coefficients. If the sampling of the underlying signal is too sparse, we will get aliasing of higher frequency modes. In flat spaces, Nyquist's theorem provides a clear prescription for uniform sampling on a grid. Since there is no such natural sampling on a sphere, the situation is more complex and depends on the pixelization scheme used \cite{Driscoll:1994:CFT:184069.184073,DBLP:journals/corr/abs-1110-6297}. Nevertheless, clear heuristic prescriptions exist. Most importantly, the typical pixel separation needs to be smaller than the wavelength of the highest frequency mode present in the underlying signal. As a natural consequence, the smoothing of the map must always precede any downgrade in resolution (or equivalently downsampling) if it is necessary to do so for numerical expediency. Similarly, any mask should be applied at the highest possible resolution and the resulting map smoothed before being down-graded.

        For the same reasons, even if the signal is correctly sampled in a band-limited sense, one must always estimate power over the entire power band, even if one is not interested in certain region of $\ell$ space. Otherwise, the power spectrum estimator will try to ``explain'' the variance associated with $\ell$ modes not considered by artificially inflating the power in the estimated modes. For the {\tt HEALPix} pixelization scheme \citep{2005ApJ...622..759G}, commonly used in cosmology and also in this paper, \cite{2013MNRAS.435.1857L} suggest the following prescription. If one wants to measure power up to $\ell_{\rm max}$, one should employ pixelisation of \texttt{Nside}=$\ell_{\rm max}/2$ and estimate band-powers up to $2 \ell_{\rm max}$, but then discard measurements above $\ell_{\rm max}$ after correcting for pixel-response suppression and  accounting for their contribution to the covariance matrix of measurements.

@@

  \section{Flat skies and Fourier transforms}\label{app:flat}
    \subsection{Spin-\texorpdfstring{$s$}{s} fields in flat sky}\label{app:flat.fields}
      In the flat sky we will label a position in the plane by two coordinates $(x,y)\equiv{\bf x}$. These coordinates can be directly related to increments in the spherical coordinates $(\theta,\varphi)$ by considering a sufficiently small map centered around the equator (i.e. $\sin\theta\sim 1$). In this case we identify the $x$ coordinate with a latitude shift ($\delta\theta=-\delta x$) and the $y$ coordinate with longitude
      shifts ($\delta\varphi=\delta y$) the differential operator $\eth$ now takes the form:
      \begin{equation}
        \eth= (\partial_x-i\partial_y),\hspace{12pt}\bar{\eth}= (\partial_x+i\partial_y).
      \end{equation}
      The analog of the standard spherical harmonics for flat skies are plane waves $e^{i{\bf k}{\bf x}}$, and the action of $\eth$ on these is:
      \begin{equation}
        \eth^s e^{i{\bf k}{\bf x}}=(ik)^se^{-i\,s\varphi_k}e^{i{\bf k}{\bf x}},\hspace{12pt}
        \bar{\eth}e^{i{\bf k}{\bf x}}=(ik)^se^{i\,s\varphi_k}e^{i{\bf k}{\bf x}}
      \end{equation}

>>>>

          \item {\bf Signal:} we generate maps of the galaxy overdensity $\delta$ and of the shear field $(\gamma_1,\gamma_2)$ for a redshift bin centred at $z\simeq1$ with a width $\Delta z\sim0.1$ and Gaussian photo-$z$ tails with width $\sigma_z=0.06$. For $\delta_g$ we assume unit bias $b(z)=1$ and no contributions from redshift-space distortions or magnification. For $\gamma$, we assume no contribution from intrinsic alignments. The maps are generated as Gaussian realizations of the corresponding angular power spectra computed with the Core Cosmology Library\footnote{\url{https://github.com/LSSTDESC/CCL}} for cosmological parameters compatible with the 2015 Planck measurements~\citet{2016A&A...594A..13P}. These power spectra are shown in Fig. \ref{fig:lss_cls}. Sample realizations of $\delta$ and $\gamma$ can be seen in Figures \ref{fig:lss_mocks_sph} and \ref{fig:lss_mocks_flat} for the curved-sky and flat-sky cases respectively.
          \item {\bf Mask:} for the curved-sky data, we use a mask built as a combination of the sky coverage output by the LSST {\tt OpSim} database~\citep{2014SPIE.9150E..15D} and a more conservative galactic cut based on the reddening map of \citet{1998wfsc.conf..297S}. To add an extra layer of complexity, we drill 100 additional 1-degree holes on the resulting combined mask. We also explore the improvement stemming from the down-weighting of higher-noise in an inverse-variance way in the presence of inhomogeneous noise (see description below), which is the reason why the default version of this mask is not binary, but traces the inverse of the noise variance (see top left panel of Fig. \ref{fig:lss_mocks_sph}). For the flat-sky data, we use a high-resolution mask constructed from the bright-object mask used for the first public data release from the HSC collaboration~\citep{2018PASJ...70S...4A}. The chosen footprint corresponds to the VVDS field, covers approximately 20 deg$^2$ and contains structure on a wide range of scales. As can be seen in the top-left panel of Figure~\ref{fig:lss_mocks_flat}, the mask is not binary, and its value at each pixel correspond to the pixel's fractional masked area.

@@

    \subsection{Curved-sky large-scale structure validation}\label{ssec:validation.lss_full}
      \begin{figure*}
        \centering
        \includegraphics[width=0.99\columnwidth]{val_cl_lss_sph}
        \includegraphics[width=0.99\columnwidth]{val_covar_lss_sph}
        \caption{{\sl Left panel:} average over all simulations in our curved-sky LSS validation suite of the six power spectra between $\delta$, $\gamma_E$ and $\gamma_B$ (solid lines). The analytical prediction for the input signal power spectra are shown as darker dashed lines, and lie mostly under the solid lines (for the non-zero cases $\delta\times\delta$, $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$). The null power spectra ($\delta\times\gamma_B$, $\gamma_E\times\gamma_B$ and $\gamma_B\times\gamma_B$) are compatible with zero. The middle subplot shows the mean residuals with respect to the input power spectra normalized by the 1$\sigma$ error over the mean (i.e. the scatter over simulations divided by $\sqrt{N_{\rm sim}}$). The residuals are scattered around zero and their fluctuations are well within those expected under the assumption of Gaussian statistics given 1000 independent realizations (see Fig. \ref{fig:lss_val_chi2}). For comparison, the lower panel show the mean relative difference for the power spectra of contaminated maps without contaminant deprojection. {\sl Right panel:} correlation matrix ($C_{ij}/\sqrt{C_{ii}C_{jj}}$) of all measured power spectra estimated from 1000 realizations.} \label{fig:lss_val_cls}
      \end{figure*}
      \begin{figure*}
        \centering
        \includegraphics[width=0.80\textwidth]{val_chi2_lss_sph}
        \caption{Distribution of the $\chi^2$ values found for each of the 1000 simulations in our curved-sky LSS validation suite (blue histograms). Results are shown for all possible cross-correlations, labeled in the top right corner of each subplot. The simulated distributions are compared with the expected $\chi^2$ distribution (solid lines) assuming Gaussian statistics for $N_{\rm dof}=681$ degrees of freedom (denoted by the dashed vertical lines). The dot-dashed vertical lines mark the $\chi^2$ value found for the mean residual with respect to the input power spectrum (see middle panel of the left panel in Fig. \ref{fig:lss_val_cls}). The associated PTE values are $\gtrsim0.1$, and thus no statistically significant bias is found by our validation suite.} \label{fig:lss_val_chi2}
      \end{figure*}
      We generate 1000 realizations of contaminated maps of the galaxy overdensity $\delta$ and cosmic shear $\gamma$ using the models described in Section \ref{sssec:validation.suite.lss} with resolution $N_{\rm side}=1024$. In our fiducial run, we included all contaminants, and used a non-binary mask that traces the shot-noise fluctuations due to the presence of stars and depth variations. For each simulation, we use \nmt~to compute all six possible cross-correlations between the $\delta$ and the $E$ and $B$ modes of $\gamma$ (which we will label $\gamma_{E,B}$), including map-level contaminant deprojection and removal of the noise and deprojection bias in the power spectrum. We use band-powers defined by averaging over groups of $\Delta\ell=3$ multipoles. The typical running time for one realization is $\sim 200$s using 32 \texttt{OpenMP} threads. Subsequent runs are typically faster since we cache the coupling matrices, which are the slowest step in the pipeline. More details about timing and scalability of \nmt~can be found in Appendix \ref{app:scalability}.

@@

      The above figure was created using $B$-mode purification. Disabling this option has no impact on the bias of the estimates, but increases its variance as we show in the Figure \ref{fig:cmb_val_sigma}. We plot diagonal error bars with (red) and without (blue) $B$-mode purification for the $BB$ (solid), $EB$ (dashed) and $EE$ (dot-dashed) power spectra. Although there is no practical improvement in the error bars for the $EE$ spectrum (there is even a slight increase due to the loss of the ambiguous $E$ modes to purification), and the improvement is only mild for the $EB$ case, very significant gains can be obtained for the $BB$ power spectrum, reducing uncertainties by up to a factor of $\sim10$ on large scales.
      \begin{figure*}
        \centering
        \includegraphics[width=0.99\columnwidth]{val_cl_lss_flat}
        \includegraphics[width=0.99\columnwidth]{val_cl_cmb_flat}
        \caption{{\sl Left and Right:} the analogues of the left panels of Figures \ref{fig:lss_val_cls} and \ref{fig:cmb_cls} for the flat-sky LSS and CMB validation sets respectively.} \label{fig:val_cls_flat}
      \end{figure*}

@@

      \subsection{Pixelization and signal band-limits}\label{app:shts.blim}
        SHT transforms as defined in Eq.~\ref{eq:sht1} are, strictly speaking, defined for continuous fields on the sky and results in an infinite number of spherical harmonic coefficients. In practice, we have to always deal with discretized sky maps with a finite number of pixels at which the field is sampled, and consequently use a finite number of spherical coefficients. If the sampling of the underlying signal is too sparse, we will get aliasing of higher frequency modes. In flat spaces, Nyquist's theorem provides a clear prescription for uniform sampling on a grid. Since there is no such natural sampling on a sphere, the situation is more complex and depends on the pixelization scheme used \citep{Driscoll:1994:CFT:184069.184073,DBLP:journals/corr/abs-1110-6297}. Nevertheless, clear heuristic prescriptions exist. Most importantly, the typical pixel separation needs to be smaller than the wavelength of the highest frequency mode present in the underlying signal. As a natural consequence, the smoothing of the map must always precede any downgrade in resolution (or equivalently downsampling) if it is necessary to do so for numerical expediency. Similarly, any mask should be applied at the highest possible resolution and the resulting map smoothed before being down-graded.

        For the same reasons, even if the signal is correctly sampled in a band-limited sense, one must always estimate power over the entire power band, even if one is not interested in certain region of $\ell$ space. Otherwise, the power spectrum estimator will try to ``explain'' the variance associated with $\ell$ modes not considered by artificially inflating the power in the estimated modes. For the {\tt HEALPix} pixelization scheme \citep{2005ApJ...622..759G}, commonly used in cosmology and also in this paper, \cite{2013MNRAS.435.1857L} suggest the following prescription: if one wants to measure power up to $\ell_{\rm max}$, one should employ pixelisation of \texttt{Nside}=$\ell_{\rm max}/2$ and estimate band-powers up to $2 \ell_{\rm max}$, but then discard measurements above $\ell_{\rm max}$ after correcting for pixel-response suppression and accounting for their contribution to the covariance matrix of measurements.

@@

  \section{Flat skies and Fourier transforms}\label{app:flat}
    \subsection{Spin-\texorpdfstring{$s$}{s} fields in flat sky}\label{app:flat.fields}
      In the flat sky we will label a position in the plane by two coordinates $(x,y)\equiv{\bf x}$. These coordinates can be directly related to increments in the spherical coordinates $(\theta,\varphi)$ by considering a sufficiently small map centered around the equator (i.e. $\sin\theta\sim 1$). In this case we identify the $x$ coordinate with a latitude shift ($\delta\theta=-\delta x$) and the $y$ coordinate with longitude shifts ($\delta\varphi=\delta y$). The differential operator $\eth$ now takes the form:
      \begin{equation}
        \eth= (\partial_x-i\partial_y),\hspace{12pt}\bar{\eth}= (\partial_x+i\partial_y).
      \end{equation}
      The analog of the standard spherical harmonics for flat skies are plane waves $e^{i{\bf k}{\bf x}}$, and the action of $\eth$ on these is:
      \begin{equation}
        \eth^s e^{i{\bf k}{\bf x}}=(ik)^se^{-i\,s\varphi_k}e^{i{\bf k}{\bf x}},\hspace{12pt}
        \bar{\eth}e^{i{\bf k}{\bf x}}=(ik)^se^{i\,s\varphi_k}e^{i{\bf k}{\bf x}}
      \end{equation}
